Irt plot extra

### 2022-03-28 manuscript.Rmd
Typo Corrected. All Supplemenatry materials are merged into one.

author: 
  - name          : "Mushfiqul Anwar Siraji"
    affiliation   : "1, *"
    email         : "mushfiqul.anwarsiraji@monash.edu"
    role:         
      - Formal Analysis
      - Visualization
      - Writing – original draft
      - Writing – review & editing;

  - name          : "Rafael Robert Lazar"
    affiliation   : "2, 3, *"
    role:
      - Data curation
      - Investigation
      - Project administration
      - Visualization
      - Writing – original draft
      - Writing – review & editing;

@@

    missing = "no"
    ) %>% add_overall() %>% bold_labels() %>% 
  # add_p() %>% add_q() %>%
modify_header(label ~ "**Variable**") %>% 
  # separate_p_footnotes() %>%
  modify_caption("Demographic Characteristics of Partinipants (n=690)") -> desc_table

@@

Figure \@ref(fig:FlowchartFig) summarizes the steps we followed while developing LEBA. In our analysis we used R statistical tool [@R-base]. **First**, we  developed a item pool of 48 items with six-point Likert type response format (0-Does not apply/I don't know, 1-Never, 2-Rarely 3-Sometimes, 4-Often, 5-Always) for our initial scale. Our purpose was to capture light exposure related behaviour. In that context, the first two response options: "Does not apply/I don't know" and "Never" were providing similar information. As such we collapsed them into one, making it a 5 point Likert type response format.

**Second**, for data collection we conducted a large scale survey. **Third**, with the data collected on the first round of the large scale survey (EFA sample; n=`r nrow(EFA.descriptives)`) we conducted descriptive and item analysis and proceeded to the exploratory factor analysis (EFA). Prior to the EFA, necessary assumptions, including sample adequacy, normality assumptions, quality of correlation matrix were assessed. Our data violated both the univariate and multivariate normality assumptions. Due to these violations and the ordinal nature of our response data, in EFA we used  polychoric correlation matrix and employed principal axis (PA) as  factor extraction method [@desjardinsHandbookEducationalMeasurement2018;@watkinsStepbyStepGuideExploratory2020]. We used a combination of factor identification method including Scree plot [@cattellScreeTestNumber1966], minimum average partials method [@velicerDeterminingNumberComponents1976], and Hull method [@lorenzo-sevaHullMethodSelecting2011] to identify factor numbers. To determine the latent structure, we followed the common guidelines : (i) no factors with fewer than three items (ii) no factors with a factor loading <0.3 (iii) no items with cross-loading > .3 across factors [@bandalosFactorAnalysisExploratory2018]. We also conducted a EFA on non-merged response options data (**Supplementary File 1**)

For reliability estimation we used internal consistency reliability coefficient ordinal $\alpha$. Though Cronbach's alpha coefficient is widely used for estimating internal consistency, it has a tendency to deflate the estimates for Likert-type data since the calculation is based on pearson-correlation matrix which requires response data to be continuous of nature [@gadermann2012estimating; @zumbo2007ordinal]. Subsequently to get  better estimates of reliability  we reported ordinal alpha for each factors using polychoric-correlation matrix [@zumbo2007ordinal]. We also estimated the internal consistency reliability of the total scale using McDonald's $\omega_t$  coefficient which is a better reliability estimate for multidimensional constructs [@dunn2014alpha; @sijtsma2009use]. Both ordinal alpha and McDonald's $\omega_t$ coefficient value range from 0 to 1 and higher value represents better reliability.

To validated the latent structure obtained in EFA, We conducted a categorical confirmatory factor analysis (CFA) with weighted least square with mean and variance adjusted (WLSMV) estimator  [@desjardinsHandbookEducationalMeasurement2018] on the data collected n our second round of data collection (CFA sample;n=`r nrow(CFA.descriptives)`). We assessed the model fit using common model fit guidelines: (i) $\chi^2$ test statistics: a non-significant test statistics is required to accept the model (ii) comparative fit index (CFI) and Tucker Lewis index (TLI):  close to .95 or above/ between 90-95 and above (iii) root mean square error of approximation (RMSEA): close to .06 or below, (iv) Standardized root mean square (SRMR): close to .08 or below [@huCutoffCriteriaFit1999; @schumacker2004beginner]. However, the $\chi^2$ test is sensitive to sample size [@brownConfirmatoryFactorAnalysis2015] and SRMR does not work well with ordinal data [@RN1272] As such, we judged the model fit using CFI, TLI, SRMR and RMSEA.

We assessed the measurement invariance (MI) of our construct between native English speakers (n=`r invariance.data.descriptives.count[1,2]`) and non native English speakers (n=`r invariance.data.descriptives.count[2,2]`) among the CFA sample (n=`r nrow(CFA.descriptives)`). MI evaluates whether a construct has the psychometric equivalence and same meaning across groups [@putnick2016measurement;@klinePrinciplesPracticeStructural2015]. We used  structural equation modelling framework to assess the measurement invariance.  We successively compared four nested models:  configural, metric, scalar, and residual models using the $\chi^2$ difference test ($\Delta \chi^2$). Among MI models, configural model is the least restrictive model and residual model is the most restrictive model. A non-significant $\Delta \chi^2$ test between two nested measurement invariance models indicates mode fit does not significantly decrease for the superior model [@dimitrov2010testing] thus allowing the superior invariance model to be accepted. [@widaman1997exploring].

@@

**Fifth**,  we developed a short form of LEBA using Item Response Theory (IRT) based analysis. We fitted each factor of LEBA using the graded response model [@samejima1997handbook] to the combined EFA and CFA sample (n =`r nrow(sem.data)`). IRT assesses the item quality by estimating item discrimination, item difficulty, item information, and test information [@bakerBasicsItemResponse2017]. Item discrimination indicates the pattern of variation in the categorical responses with the changes in latent trait level ($\theta$). Item information curve (IIC) indicates the amount of information an item carries along the latent trait continuum. Here, we reported  the item discrimination parameter and categorize the items according to the suggestions of @bakerBasicsItemResponse2017 : none = 0; very low =0.01 to 0.34; low = 0.35 to 0.64; moderate = 0.65 to 1.34 ; high = 1.35 to 1.69; very high >1.70. We discarded the items with relatively flat item information curve (information <.2) to develop the short form of LEBA. We also assessed the precision of the short LEBA using Test information curve (TIC). TIC  indicates the amount of information an the full-scale carry along the latent trait continuum. 
Item fit and person fit of the fitted IRT models were also analyzed to gather more evidence on validity and meaningfulness of our Tool [@desjardinsHandbookEducationalMeasurement2018]. Item fit was evaluated using  the RMSEA value obtained from Signed-$\chi^2$ index implementation, RMSEA value $\le$.06 was considered adequate item fit. Person fit was estimated using standardized fit index Zh statistics [@drasgow1985appropriateness]. Zh < -2 was be considered as a misfit [@drasgow1985appropriateness].

@@

After reviewing the literature, we identified several light exposure related scale. However, no scales specifically measuring the behavioural component of light exposure were found (**Supplementary Table 1**). As such, all authors in collaboration of an expert panel developed a comprehensive item pool of 48 items. The expert panel composed of all authors and researchers from the fields of chronobiology, light research, neuroscience and psychology. The 48 items were then judged  based on their relevance and representativeness of the construct "Light Exposure Related Behaviour" by the expert panel. The panel members independently judged each of the items in terms of their relevance and representativeness and suggested required modification, if there is any. The author team acknowledged the suggestions and amended the items as required thus creating a 48-item scale.

@@

**Supplementary Figure 1** summarizes the univariate descriptive statistics for the 48 items among  EFA sample (n =`r nrow(EFA.data)`.  Our data violated both univariate normality [@shapiroAnalysisVarianceTest1965] and multivariate normality assumptions [@mardiaMeasuresMultivariateSkewness1970]. Multivariate skew was  `r printnum(mardia$b1p)` (p <0.001) and multivariate kurtosis was `r printnum(mardia$b2p)` (p <0.001). Due to these violations and ordinal nature of the response data polychoric correlations over Pearson's correlations was chosen [@desjardinsHandbookEducationalMeasurement2018]. The corrected item-total correlation ranges between `r apa(low.r.corec,2,F)` -`r apa(high.r.corec,2,F)`. However, no item was discarded based on descriptive statistics or item analysis.

@@

Sampling adequacy was checked using Kaiser-Meyer-Olkin (KMO) measures of sampling adequacy using the EFA sample (n =`r nrow(EFA.data)` [@kaiserIndexFactorialSimplicity1974] . The overall KMO vale for 48 items was `r printnum(KMO$MSA)` which was above the cutoff value (.50) indicating adequate sample size [@hutchesonMultivariateSocialScientist1999].  Bartlett's test of sphericity [@bartlettNoteMultiplyingFactors1954], $\chi^2$ (`r bartlet$df`) = `r apa(bartlet$chisq, 2, T)`, p < .001 indicated the correlations between items are adequate for conducting the EFA. However only `r apa(cor.per, 2,T)`% of the inter-item correlation coefficients were greater than |.30|. The inter-item correlation coefficients ranged between `r min.cor` to `r max.cor`. Figure \@ref(fig:efa-plot-print)A depicts the correlation matrix.

@@

Scree plot ( Figure \@ref(fig:efa-plot-print)B) suggested a six-factor solution. However, the minimum average partial (MAP) [@velicerDeterminingNumberComponents1976] method (**Supplementary Table 3**) and Hull method [@lorenzo-sevaHullMethodSelecting2011] ( Figure \@ref(fig:efa-plot-print)C) suggested a five-factor solution. As a result, we tested both five-factor and six-factor solutions.

@@

```{r efa-plot-print, fig.align= 'center', fig.cap='(ref:efa-plot-print)',  warning=FALSE, results= 'asis', out.height='90%', out.width='110%'}

@@

With the initial 48 items we conducted three rounds of EFA and gradually discarded problematic items. (cross-loading items and poor factor loading (<.30) items). Finally, a five-factor EFA solution with 25 items was accepted with all factor-loading higher than .30 and no cross-loading greater than .30. Table \@ref(tab:EFATable) displays the factor-loading (structural coefficients) and communality of the items. The absolute value of the factor-loading ranged from `r apa(min.loadings,2,F)` to `r apa(max.loadings, 2, F)` indicating strong coefficients. The commonalities ranged between `r apa(min.com,2,F)` to `r apa(max.com, 2, F)`. However, the histogram of the absolute values of non-redundant residual-correlations (Figure \@ref(fig:efa-plot-print)D) showed `r apa(sumR,0, T)`% correlations were greater than the absolute value of .05, indicating a possible under-factoring. [@desjardinsHandbookEducationalMeasurement2018]. Subsequently, we fitted a six-factor solution. However, a factor emerged with only two salient variables, thus disqualifying the six-factor solution (**Supplementary Table 4**).

@@

**Supplementary Table 6** summarizes the item fit indexes of the 18 items. All of the items had RMSEA value $\le$.06 indicating adequate fit of the items to the fitted IRT model. **Supplementary Figure 4** depicts the person fit Zh statistics histogram of out fitted models. Zh statistics are larger than -2 for most participants, suggesting a good person fit of the selected IRT models.

>>>>

author: 
  - name          : "Mushfiqul Anwar Siraji"
    affiliation   : "1, *"
    email         : "mushfiqul.anwarsiraji@monash.edu"
    role:         
      - Formal Analysis
      - Visualization
      - Writing – original draft
      - Writing – review & editing

  - name          : "Rafael Robert Lazar"
    affiliation   : "2, 3, *"
    role:
      - Data curation
      - Investigation
      - Project administration
      - Visualization
      - Writing – original draft
      - Writing – review & editing

@@

    missing = "no"
    ) %>% add_overall() %>% bold_labels() %>% 
  # add_p() %>% add_q() %>%
modify_header(label ~ "**Variable**") %>% 
  # separate_p_footnotes() %>%
  modify_caption("Demographic Characteristics of Participants (n=690)") -> desc_table

@@

Figure \@ref(fig:FlowchartFig) summarizes the steps we followed while developing LEBA. In our analysis we used R statistical tool [@R-base]. **First**, we  developed a item pool of 48 items with six-point Likert type response format (0-Does not apply/I don't know, 1-Never, 2-Rarely 3-Sometimes, 4-Often, 5-Always) for our initial scale. Our purpose was to capture light exposure related behaviour. In that context, the first two response options: "Does not apply/I don't know" and "Never" were providing similar information. As such we collapsed them into one, making it a 5 point Likert type response format (1-Never, 2-Rarely 3-Sometimes, 4-Often, 5-Always).

**Second**, for data collection we conducted a large scale survey. **Third**, with the data collected on the first round of the large scale survey (EFA sample; n=`r nrow(EFA.descriptives)`) we conducted descriptive and item analysis and proceeded to the exploratory factor analysis (EFA). Prior to the EFA, necessary assumptions, including sample adequacy, normality assumptions, quality of correlation matrix were assessed. Our data violated both the univariate and multivariate normality assumptions. Due to these violations and the ordinal nature of our response data, in EFA we used  polychoric correlation matrix and employed principal axis (PA) as  factor extraction method [@desjardinsHandbookEducationalMeasurement2018;@watkinsStepbyStepGuideExploratory2020]. We used a combination of factor identification method including Scree plot [@cattellScreeTestNumber1966], minimum average partials method [@velicerDeterminingNumberComponents1976], and Hull method [@lorenzo-sevaHullMethodSelecting2011] to identify factor numbers. To determine the latent structure, we followed the common guidelines : (i) no factors with fewer than three items (ii) no factors with a factor loading <0.3 (iii) no items with cross-loading > .3 across factors [@bandalosFactorAnalysisExploratory2018]. 
<!-- We also conducted an EFA on non-merged response options data set (**Supplementary File 1**). -->

Though Cronbach's internal consistency coefficient alpha is widely used for estimating internal consistency, it has a tendency to deflate the estimates for Likert-type data since the calculation is based on pearson-correlation matrix which requires response data to be continuous in nature [@gadermann2012estimating; @zumbo2007ordinal]. Subsequently to get  better estimates of reliability we reported ordinal alpha for each factors obtained in EFA using polychoric-correlation matrix [@zumbo2007ordinal]. We also estimated the internal consistency reliability of the total scale using McDonald's $\omega_t$  coefficient which is a better reliability estimate for multidimensional constructs [@dunn2014alpha; @sijtsma2009use]. Both ordinal alpha and McDonald's $\omega_t$ coefficient value range from 0 to 1 and higher value represents better reliability.

To validated the latent structure obtained in EFA, We conducted a categorical confirmatory factor analysis (CFA) with weighted least square with mean and variance adjusted (WLSMV) estimator  [@desjardinsHandbookEducationalMeasurement2018] on the data collected in the second round of data collection (CFA sample; n=`r nrow(CFA.descriptives)`). We assessed the model fit using common model fit guidelines: (i) $\chi^2$ test statistics: a non-significant test statistics is required to accept the model (ii) comparative fit index (CFI) and Tucker Lewis index (TLI):  close to .95 or above/ between .90-.95 and above (iii) root mean square error of approximation (RMSEA): close to .06 or below, (iv) Standardized root mean square (SRMR): close to .08 or below [@huCutoffCriteriaFit1999; @schumacker2004beginner]. However, the $\chi^2$ test is sensitive to sample size [@brownConfirmatoryFactorAnalysis2015] and SRMR does not work well with ordinal data [@RN1272] As such, we judged the model fit using CFI, TLI and RMSEA.

We assessed the measurement invariance (MI) of our scale between native English speakers (n=`r invariance.data.descriptives.count[1,2]`) and non native English speakers (n=`r invariance.data.descriptives.count[2,2]`) in the CFA sample (n=`r nrow(CFA.descriptives)`). MI evaluates whether a construct has the psychometric equivalence and same meaning across groups [@putnick2016measurement;@klinePrinciplesPracticeStructural2015]. We used  structural equation modelling framework to assess the measurement invariance.  We successively compared four nested models:  configural, metric, scalar, and residual models using the $\chi^2$ difference test ($\Delta \chi^2$). Among MI models, configural model is the least restrictive model and residual model is the most restrictive model. A non-significant $\Delta \chi^2$ test between two nested measurement invariance models indicates mode fit does not significantly decrease for the superior model thus allowing the superior invariance model to be accepted [@dimitrov2010testing; @widaman1997exploring].

@@

**Fifth**,  we developed a short form of LEBA using Item Response Theory (IRT) based analysis. We fitted each factor of LEBA using the graded response model [@samejima1997handbook] to the combined EFA and CFA sample (n =`r nrow(sem.data)`). IRT assesses the item quality by estimating item discrimination, item difficulty, item information curve, and test information curve [@bakerBasicsItemResponse2017]. Item discrimination indicates how well a particular item can differentiate between participants across the given latent trait continuum ($\theta$). Item difficulty corresponds to the latent trait level at which the probability of endorsing a particular response option is 50%. Item information curve (IIC) indicates the amount of information an item carries along the latent trait continuum. Here, we reported  the item discrimination parameter and categorize the items according to the suggestions of @bakerBasicsItemResponse2017 : none = 0; very low =0.01 to 0.34; low = 0.35 to 0.64; moderate = 0.65 to 1.34 ; high = 1.35 to 1.69; very high >1.70. We discarded the items with relatively flat item information curve (information <.2) to develop the short form of LEBA. We also assessed the precision of the short LEBA using Test information curve (TIC). TIC  indicates the amount of information an the full-scale carry along the latent trait continuum. 
Item fit and person fit of the fitted IRT models were also analyzed to gather more evidence on validity and meaningfulness of our Tool [@desjardinsHandbookEducationalMeasurement2018]. Item fit was evaluated using  the RMSEA value obtained from Signed-$\chi^2$ index implementation, RMSEA value $\le$.06 was considered adequate item fit. Person fit was estimated using standardized fit index Zh statistics [@drasgow1985appropriateness]. Zh < -2 was considered as a misfit [@drasgow1985appropriateness].

@@

After reviewing the literature, we identified several light exposure related scale. However, no scale specifically measuring the behavioural component of light exposure were found (**Supplementary Table 1**). As such, an expert panel developed a comprehensive item pool of 48 items. The expert panel composed of all authors and researchers from the fields of chronobiology, light research, neuroscience and psychology. The 48 items were then judged  based on their relevance and representativeness of the construct "Light Exposure Related Behaviour" by the expert panel. The panel members independently judged each of the items in terms of their relevance and representativeness and suggested required modification, if there is any. The author team acknowledged the suggestions and amended the items as required thus creating a 48-item scale.

@@

**Supplementary Figure 1** summarizes the univariate descriptive statistics for the 48 items among  EFA sample (n =`r nrow(EFA.data)`.  Our data violated both univariate normality [@shapiroAnalysisVarianceTest1965] and multivariate normality assumptions [@mardiaMeasuresMultivariateSkewness1970]. Multivariate skew was  `r printnum(mardia$b1p)` (p <0.001) and multivariate kurtosis was `r printnum(mardia$b2p)` (p <0.001). The corrected item-total correlation ranges between `r apa(low.r.corec,2,F)` -`r apa(high.r.corec,2,F)`. However, no item was discarded based on descriptive statistics or item analysis.

@@

Sampling adequacy was checked using Kaiser-Meyer-Olkin (KMO) measures of sampling adequacy using the EFA sample (n =`r nrow(EFA.data)`) [@kaiserIndexFactorialSimplicity1974] . The overall KMO vale for 48 items was `r printnum(KMO$MSA)` which was above the cutoff value (.50) indicating adequate sample size [@hutchesonMultivariateSocialScientist1999].  Bartlett's test of sphericity [@bartlettNoteMultiplyingFactors1954], $\chi^2$ (`r bartlet$df`) = `r apa(bartlet$chisq, 2, T)`, p < .001 indicated the correlations between items are adequate for conducting the EFA. However only `r apa(cor.per, 2,T)`% of the inter-item correlation coefficients were greater than |.30|. The inter-item correlation coefficients ranged between `r min.cor` to `r max.cor`. Figure \@ref(fig:efa-plot-print)-A depicts the correlation matrix.

@@

Scree plot ( Figure \@ref(fig:efa-plot-print)-B) suggested a six-factor solution. However, the minimum average partial (MAP) [@velicerDeterminingNumberComponents1976] method (**Supplementary Table 3**) and Hull method [@lorenzo-sevaHullMethodSelecting2011] ( Figure \@ref(fig:efa-plot-print)C) suggested a five-factor solution. As a result, we tested both five-factor and six-factor solutions.

@@

```{r efa-plot-print, fig.align= 'center', fig.cap='(ref:efa-plot-print)',  warning=FALSE,  out.height='60%', out.width='110%'}

@@

With the initial 48 items we conducted three rounds of EFA and gradually discarded problematic items (cross-loading items and poor factor loading (<.30) items). Finally, a five-factor EFA solution with 25 items was accepted with all factor-loading higher than .30 and no cross-loading greater than .30. Table \@ref(tab:EFATable) displays the factor-loading (structural coefficients) and communality of the items. The absolute value of the factor-loading ranged from `r apa(min.loadings,2,F)` to `r apa(max.loadings, 2, F)` indicating strong coefficients. The commonalities ranged between `r apa(min.com,2,F)` to `r apa(max.com, 2, F)`. However, the histogram of the absolute values of non-redundant residual-correlations (Figure \@ref(fig:efa-plot-print)D) showed `r apa(sumR,0, T)`% correlations were greater than the absolute value of .05, indicating a possible under-factoring. [@desjardinsHandbookEducationalMeasurement2018]. Subsequently, we fitted a six-factor solution. However, a factor emerged with only two salient variables, thus disqualifying the six-factor solution (**Supplementary Table 4**).

@@

**Supplementary Table 6** summarizes the item fit indexes of the 18 items. All of the items had RMSEA value $\le$.06 indicating adequate fit of the items to the fitted IRT model. **Supplementary Figure 4** depicts the person fit Zh statistics histogram of our fitted models. Zh statistics are larger than -2 for most participants, suggesting a good person fit of the selected IRT models.
